Phonetic Transcription by Untrained Annotators

### 2018-11-21 2018-clarin-pisa/makon-data.tex
2018 papers

\author{First Author \\
  Department (optional)\\
  University of City, Country \\
  {\tt email@domain} \\\And % if needed: this makes a second column
  Second Author \\
  Department (optional)\\
  University Name without city \\
  City, Country \\
 {\tt email@domain} \\
% \AND % if needed: this makes a second row
%  Third Author \\
%  Department (optional)\\
%  University of City, Country \\
%  {\tt email@domain} \\\And
%  Fourth Author \\
%  Department (optional)\\
%  University Name without city \\
%  City, Country \\
%  {\tt email@domain} \\
}

@@

>>>>

\author{Old\v{r}ich Kr\r{u}za \\
  Institute of Formal and Applied Linguistics\\
  Faculty of Mathematics and Physics, Charles University in Prague, Czech
  Republic \\
  {\tt kruza@ufal.mff.cuni.cz} %\\\And % if needed: this makes a second column
%  Second Author \\
%  Department (optional)\\
%  University Name without city \\
%  City, Country \\
% {\tt email@domain} \\
% \AND % if needed: this makes a second row
%  Third Author \\
%  Department (optional)\\
%  University of City, Country \\
%  {\tt email@domain} \\\And
%  Fourth Author \\
%  Department (optional)\\
%  University Name without city \\
%  City, Country \\
%  {\tt email@domain} \\
}

@@

\section*{Acknowledgments}

The research was supported by SVV project number 260 453.\\
\\
This work has been using language resources stored
and distributed  by the LINDAT/CLARIN project of the Ministry of Education,
Youth and Sports of the Czech Republic (project LM2015071).

### 2018-11-21 2018-ftc-vancouver/webapp.tex
2018 papers

Mako\v{n}'s written works present his teachings in a systematic, comprehensible
fashion, while the recordings offer bonuses: talks tailored to the audience,
answers to questions, personal experiences, behind-the-scenes to the books etc.
The archive is freely
accessible\footnote{https://lindat.mff.cuni.cz/repository/xmlui/handle/11372/LRT-1455}
under the CC-BY license.

@@

The most common tasks have keyboard shortcuts: \texttt{ctrl+space} for
play/pause and \texttt{ctrl+enter} for submitting a correction.

@@

The efficiency of repositioning a rectangle is supported by the fact that we can
calculate the coordinates of all rendered words once and only recalculate them
in two cases: 1) In the rare event of screen resize and 2) when a corrected
segment is merged into the transcription, in which case we only need to
recalculate for the words further in the document.

@@

One of the most profound measures in ergonomy are keyboard shortcuts. The most
common task is pausing and resuming playback. Both oTranscribe and Transcribe
use the \texttt{esc} key for that, and Transcriber uses the \texttt{tab} key.
We chose \texttt{ctrl+space} combination. We argue that \texttt{esc} is not the
best of options because the distance the fingers have to travel from the
alphanumeric keys causes a noticeable delay. This can lead to missing a pause
between words. The \texttt{tab} key as chosen by Transcriber is a splendid choice
from the ergonomy point of view and there is no reason not to use it in a
dedicated user interface. However, in the browser, where the \texttt{tab} key
has as native use, re-binding it could lead to confusion and irritation. The
space bar is probably the easiest-to-find key in all situations and dedicating
\texttt{ctrl} to all application-specific commands as opposed to single keys
lends a sense of consistency, we believe.

@@

Another point of near-future endeavor is higher-level work with the contents. By
this we mean that we wouldd like to use both automatic processing methods and the
users to do semantic analysis of the talks: What topic is covered where? What
topics are covered at all? Which talks relate to which written works?, and
similar questions.

>>>>

Mako\v{n}'s written works present his teachings in a systematic, comprehensive
fashion, while the recordings offer bonuses: talks tailored to the audience,
answers to questions, personal experiences, behind-the-scenes to the books etc.
The archive is freely
accessible\footnote{https://lindat.mff.cuni.cz/repository/xmlui/handle/11372/LRT-1455}
under the CC-BY license.

@@

The commonest tasks have keyboard shortcuts: \texttt{ctrl+space} for
play/pause and \texttt{ctrl+enter} for submitting a correction.

@@

The efficiency of repositioning a rectangle is supported by the fact that we can
calculate the coordinates of all rendered words once and only recalculate them
in two cases: 1) In the rare event of screen resize and 2) when a corrected
segment is merged into the transcription, in which case we only need to
recalculate for the words further in the document.\footnote{We could even stop
the recalculation as soon as we find that the new horizontal coordinate of a
word is left untouched, and add the difference in the vertical coordinate to all
subsequent words, i.e. when a line stays the same, so do all below it.}

@@

One of the most profound measures in ergonomy are keyboard shortcuts. The most
common task is pausing and resuming playback. Both oTranscribe and Transcribe
use the \texttt{esc} key for that, and Transcriber uses the \texttt{tab} key.
We chose \texttt{ctrl+space} combination. We argue that \texttt{esc} is not the
best of options for desktops because the distance the fingers have to travel
from the alphanumeric keys causes a noticeable delay. This can lead to missing a
pause between words. The \texttt{tab} key as chosen by Transcriber is a splendid
choice from the ergonomy point of view and there is no reason not to use it in a
dedicated user interface. However, in the browser, where the \texttt{tab} key
has as native use, re-binding it could lead to confusion and irritation. The
space bar is probably the easiest-to-find key in all situations and dedicating
\texttt{ctrl} to all application-specific commands as opposed to single keys
lends a sense of consistency, we believe.

@@

Another point of near-future endeavor is higher-level work with the contents. By
this we mean that we would like to use both automatic processing methods and the
users to do semantic analysis of the talks: What topic is covered where? What
topics are covered at all? Which talks relate to which written works?, and
similar questions.

### 2018-11-21 2018-itat-krompachy/paper.tex
2018 papers

@@

We always assume an existing transcription,\footnote{The seed training data
was of magnitude in order of minutes and has been annotated using Transcriber.}
so we can see the user's contribution as a correction. Every submission has
the form of replacing a text segment with another. Since the transcriptions are
time-aligned to the audio, we also know exactly what is the corresponding audio
segment to the text submitted.

@@

False positives (when the system accepts a wrong transcription) present a
problem, since the error will enter the data set. But false negatives can often
be circumvented by submitting the transcription divided in different segments.
Of course, this method can also be used to force a wrong transcription but we
assume no malevolence on the part of the users.

Apart from catching wrong transcription, the forced alignment mechanism provides
exact synchronization on the time axis. This is a completely missing element in
the case of virtually all programs for computer-aided transcription. For some
examples, Transcriber,\footnote{http://trans.sourceforge.net/} a veteran
open-source transcribing program for Linux, expects the user to provide
alignment on the level of phrases;
Transcribe,\footnote{https://transcribe.wreally.com/} a commercial web-based
transcribing tool, allows the user to add timestamps anywhere in the text. There
is no acoustic model, hence nothing to match against.

@@

\begin{table*}[htpb]
\fontspec{DoulosSIL}
\begin{center}
\begin{tabular}{|l|l|l|l||l|l|l|l|}
\hline
IPA & PACal & common grapheme & IPA & PACal &  common grapheme \\
\hline
% TODO: IPA
a  & a   & a      &     ɱ  & mg  & tra\underline{m}vaj \\
aː & aa  & á      &     n  & n   & \underline{n}e \\
aʊ̯ & aw  & au     &     ŋ  & ng  & ta\underline{n}k \\
b  & b   & b      &     ɲ  & nj  & \v{n} \\
t͡s & c   & c      &     o  & o   & o \\
t͡ʃ & ch  & č      &     oː & oo  & ó \\
d  & d   & d      &     oʊ̯ & ow  & ou \\
ɟ  & dj  & \v{d}  &     p  & p   & p \\
d͡z & dz  & dz     &     r  & r   & r \\
d͡ʒ & dzh & dž     &     r̝̊  & rsh & t\underline{\v{r}}i \\
ɛ  & e   & e      &     r̝  & rzh & \underline{\v{r}}íz \\
ɛː & ee  & é      &     s  & s   & s \\
eʊ̯ & ew  & eu     &     ʃ  & sh  & š \\
f  & f   & f      &     t  & t   & t \\
g  & g   & g      &     c  & tj  & \v{t} \\
ɦ  & h   & h      &     ʊ  & u   & u \\
i  & i   & i      &     uː & uu  & ú, \r{u} \\
iː & ii  & í      &     v  & v   & v \\
j  & j   & j      &     x  & x   & ch \\
k  & k   & k      &     z  & z   & z \\
l  & l   & l      &     ʒ  & zh  & ž \\
m  & m   & \underline{m}ák
                  &        & sil & \\
   &     &        &        & sp  & \\
\hline
\end{tabular}
\caption{Phonemes used in transcription}\label{tab:phones}
\end{center}
\end{table*}
\normalfont

@@

The module includes two-way transduction, although only the one from array of
phonemes to human-readable phonetic respelling is needed in our application.
Still, the user can mark up special-case pronunciation with the apostrophe, like
the sequence of phonemes \texttt{c} and \texttt{h} with the string \texttt{c'h}.
The need has never occurred during the six years' lifespan of the application.

@@

\begin{table*}[htpb]
\begin{center}
\begin{tabular}{|l r|l r|l r|l r|}
\hline
Correct spelling            & \#
	& wrong phonetic pronunc.   & \#
		& correct pronunciation     & \#
			& phon. respel.             & \# \\
\hline
Moody & 2 & \texttt{m o o d i} & 0 & \texttt{m uu d i} & 4 & {\em múdy}, {\em můdy} & 2 \\
Descartes & 2 & \texttt{d e s c a r t e s} & 0 & \texttt{d e k aa r t} & 4 & {\em dekárt} & 2   \\
Weinfurter & 30 & \texttt{v e j n f u r t e r} & 13 & \texttt{v a j n f u r t r} & 19 & {\em vajnfurtr} & 2 \\
Michelangelo & 6 & \texttt{m i x e l a ng g e l o} & 2 & \texttt{m i k e l a n dzh e l o} & 4 & {\em mikelandželo} & 0 \\
\hline
\end{tabular}
\caption{Examples of non-standard pronunciation in the manually transcribed data}\label{tab:pronunc}
\end{center}
\end{table*}

\begin{table*}[htpb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline
 & phonetically correct & phonetically incorrect \\
\hline
orthographically correct & 25 & 15 \\
\hline
orthographically incorrect & 6 & 0 \\
\hline
\end{tabular}
\caption{Success rate for phonetic and orthographic representation of foreign
words}\label{tab:pronunc-rate}
\end{center}
\end{table*}

@@

We have presented an application that has been providing access to the
extensive corpus of Karel Makoň and to acquire an almost complete transcription
thereof. Nearly 70 hours corresponding to over 600,000 word forms have been
transcribed manually with minimal financial\footnote{In early stages, we kept a paid
annotator to test the application.} as well as development\footnote{The system
has been written by a single developer.} costs. Only some of the volunteer have
indulged instruction time in order of minutes. The rest of the corpus has been
transcribed using an ASR system trained on these ever-growing data.

>>>>

\subsection{Architecture Overview}

The system consists of

\begin{enumerate}
\item{
    The corpus in compressed audio format. We use mp3 and ogg/vorbis to
    accomodate most browsers.  These data are hosted on an external CDN.
}
\item{
    The exact copy of the corpus in parametrized (MFCC) format. These data
    reside on the back-end server.
}
\item{
    A complete, aligned transcription of the recordings, hosted on the back-end
    server and mirrored on a CDN.
}
\item{
    Acoustic model trained on the human-transcribed part of the corpus.
}
\item{
    Language model trained using Srilm\cite{stolcke2002srilm} on a combination
    of publicly available Czech texts, Karel Makoň's written works, and both the
    human-submitted and automatically-acquired transcription.
}
\item{
    Back-end API for collecting correcions to the transcription, serving the
    transcription and allowing full-text search with
    elasticsearch\footnote{https://www.elastic.co/products/elasticsearch}.
}
\item{
    Separately hosted front-end web application serving as an interface for
    playing the recordings, synchronously displaying the transcriptions and
    collecting the corrections from users.
}
\end{enumerate}

To get the initial transcription, we have manually transcribed some 10 minutes
of the material using Transcriber\footnote{http://trans.sourceforge.net/},
trained an acoustic model on it and recognized the whole data using it.

@@

We always assume an existing transcription, so we can see the user's
contribution as a correction. Every submission has the form of replacing a text
segment with another. Since the transcriptions are time-aligned to the audio, we
also know exactly what is the corresponding audio segment to the text submitted.

@@

False positives (when the system accepts a wrong transcription) present a
problem, since the error will enter the training data set. But users can often
circumvent false negatives by submitting the transcription divided in different
segments.  Of course, this method can also be used to force a wrong
transcription but we assume no malevolence on the part of the users.

Apart from catching wrong transcription, the forced alignment mechanism provides
exact synchronization on the time axis. This is a completely missing element in
the case of virtually all programs for computer-aided transcription. For some
examples, Transcriber, a veteran open-source transcribing program for Linux,
expects the user to provide alignment on the level of phrases;
Transcribe,\footnote{https://transcribe.wreally.com/} a commercial web-based
transcribing tool, allows the user to add timestamps anywhere in the text. There
is no acoustic model, hence nothing to match against.

@@

\begin{table*}[htpb]
\fontspec{DoulosSIL}
\begin{center}
\begin{tabular}{|l|l|l||l|l|l|}
\hline
IPA & PACal & common grapheme & IPA & PACal &  common grapheme \\
\hline
% TODO: IPA
a  & a   & a      &     ɱ  & mg  & tra\underline{m}vaj \\
aː & aa  & á      &     n  & n   & \underline{n}e \\
aʊ̯ & aw  & au     &     ŋ  & ng  & ta\underline{n}k \\
b  & b   & b      &     ɲ  & nj  & \v{n} \\
t͡s & c   & c      &     o  & o   & o \\
t͡ʃ & ch  & č      &     oː & oo  & ó \\
d  & d   & d      &     oʊ̯ & ow  & ou \\
ɟ  & dj  & \v{d}  &     p  & p   & p \\
d͡z & dz  & dz     &     r  & r   & r \\
d͡ʒ & dzh & dž     &     r̝̊  & rsh & t\underline{\v{r}}i \\
ɛ  & e   & e      &     r̝  & rzh & \underline{\v{r}}íz \\
ɛː & ee  & é      &     s  & s   & s \\
eʊ̯ & ew  & eu     &     ʃ  & sh  & š \\
f  & f   & f      &     t  & t   & t \\
g  & g   & g      &     c  & tj  & \v{t} \\
ɦ  & h   & h      &     ʊ  & u   & u \\
i  & i   & i      &     uː & uu  & ú, \r{u} \\
iː & ii  & í      &     v  & v   & v \\
j  & j   & j      &     x  & x   & ch \\
k  & k   & k      &     z  & z   & z \\
l  & l   & l      &     ʒ  & zh  & ž \\
m  & m   & \underline{m}ák
                  &        & sil & \\
   &     &        &        & sp  & \\
\hline
\end{tabular}
\caption{Phonemes used in transcription}\label{tab:phones}
\end{center}
\end{table*}
\normalfont

@@

The module includes two-way transduction, although only the one from array of
phonemes to human-readable phonetic respelling is needed in our application.
Still, the user can mark up special-case pronunciation with the apostrophe, like
the sequence of phonemes \texttt{o} and \texttt{u} with the string \texttt{o'u}.
The need has never occurred during the six years' lifespan of the application.

@@

\begin{table*}[htpb]
\begin{center}
\begin{tabular}{|l r|l r|l r|l r|}
\hline
Correct spelling            & \#
	& wrong phonetic pronunc.   & \#
		& correct pronunciation     & \#
			& phon. respel.             & \# \\
\hline
Moody & 2 & \texttt{m o o d i} & 0 & \texttt{m uu d i} & 4 & {\em múdy}, {\em můdy} & 2 \\
Descartes & 2 & \texttt{d e s c a r t e s} & 0 & \texttt{d e k aa r t} & 4 & {\em dekárt} & 2   \\
Weinfurter & 30 & \texttt{v e j n f u r t e r} & 13 & \texttt{v a j n f u r t r} & 19 & {\em vajnfurtr} & 2 \\
Michelangelo & 6 & \texttt{m i x e l a ng g e l o} & 2 & \texttt{m i k e l a n dzh e l o} & 4 & {\em mikelandželo} & 0 \\
\hline
\end{tabular}
\caption{Examples of non-standard pronunciation in the manually transcribed data}
\label{tab:pronunc}
\end{center}
\end{table*}

\begin{table*}[htpb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline
 & phonetically correct & phonetically incorrect \\
\hline
orthographically correct & 25 & 15 \\
\hline
orthographically incorrect & 6 & 0 \\
\hline
\end{tabular}
\caption{Success rate for phonetic and orthographic representation of foreign
words based on data from table\ref{tab:pronunc}}
\label{tab:pronunc-rate}
\end{center}
\end{table*}

@@

We have presented an application that has been providing access to the
extensive corpus of Karel Makoň and to acquire an almost complete transcription
thereof. Nearly 70 hours corresponding to over 600,000 word forms have been
transcribed manually with minimal financial\footnote{In early stages, we kept a paid
annotator to test the application.} as well as development\footnote{The system
has been written by a single developer.} costs. Only some of the volunteers have
indulged instruction time in order of minutes. The rest of the corpus has been
transcribed using an ASR system trained on these ever-growing data.
